Illuminating a Dark Lens : A Type Ia Supernova Magnified by Galaxy Cluster Abell 2744

### 2015-01-21 snTomas.tex
Updated detection figure (fig 1), fixed phot table

\section{Discovery and Follow-up}
\label{sec:DiscoveryAndFollowup}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/discovery_image}
\caption{  
\textcolor{red}{PLACEHOLDER IMAGE}
SN \tomas\ in the Abell 2744 field.  The main panel shows a
UV-optical-IR 3-color composite image constructed from all available
HST imaging of the Abell 2744 cluster field.  The inset panel shows
the immediate vicinity of SN \tomas, approximately 40\arcsec\ from
the center of the cluster.  This inset image is 
a combination of all ACS optical imaging from the HFF program, collected
when the SN was detectable.}
\end{center}
\end{figure*}

@@

We fit the SN \tomas\ spectrum in two steps.  First we determine a
spectral classification -- and get a preliminary estimate of the
redshift and phase -- using the SuperNova IDentification
software \citep{Blondin:2010}.  Second, we refine the redshift and
phase using a custom Type Ia spectral template matching
program.  Details of the fitting are given below and results are
summarized in Table~\ref{tab:SpectralFits}.

@@

The SNID program is designed to estimate the type, redshift, and phase
of a SN spectrum through cross-correlation matching with a library of
template spectra.  To account for possible distortions in the broad
shape of the SN pseudo-continuum due to dust or instrumental
calibration effects, SNID divides each SED by a smooth cubic spline
fit. This effectively removes the shape of the SN pseudo-continuum to
leave behind a flat SED superimposed with spectral absorption and
emission features.  It is these features which drive the
cross-correlation fit, so the SNID approach is insensitive to the
overall color of the SED.  We used v2.0 of the SNID template library,
which includes template SEDs covering all Type Ia and Core Collapse
sub-classes, and has recently been updated with corrections and
improvements to the Type Ib/c templates \citep{Liu:2014}.

Using SNID to fit the \tomas\ spectrum, we find strong evidence that
the spectral type is a normal Type Ia, with $>$97\% of the matching
SEDs coming from Type Ia spectra.  The best-fit SEDs are also
consistent in redshift at $z=1.32\pm0.01$, and in phase at $0.4\pm1.7$
rest-frame days from peak (meaning MJD$_{\rm pk}$=56813$\pm$4).  At
this redshift, the absorption at $\sim$8700\AA\ corresponds to the
blended \ion{Ca}{2} H\&K features.  This \ion{Ca}{2} absorption is
commonly seen in Type Ia SN spectra near maximum light, although it is
also prominent in the spectra of Type Ib and Ic core collapse SNe
(\CCSNe).  The red color of the \tomas\ SED is qualitatively
consistent with a redshift of $z>1$ -- although this information was
not used by SNID for the template matching.  As we will see in
Section~\ref{sec:LightCurve}, this spectral classification of
SN \tomas\ is supported by the photometric information, which is also
most consistent with a Type Ia SN at $z=1.3$.

\subsection{Fitting with Additional Type Ia Templates}
\label{sec:FittingWithAdditionalTypeIaTemplates}

@@

>>>>

\section{Discovery, Follow-up, and Data Processing}
\label{sec:DiscoveryAndFollowup}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/discovery_image}
\caption{  
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.\footnote{Image Credit: NASA,
ESA, and J. Lotz, M. Mountain, A. Koekemoer, and the HFF Team
(STScI)}  The inset panels at right show F814W imaging of the
immediate vicinity of SN \tomas, approximately 40\arcsec\ from the
center of the cluster.  The top panel shows the template image,
combining all data prior to the SN appearance.  Labeled ellipses mark the nearest galaxies and their redshift constraints: a photometric redshift for the most likely host galaxy is marked in blue, and the spectroscopic redshift for a likely background galaxy is given in yellow. The bottom panel is
constructed from all HFF F814W imaging taken while the SN was detectable, and marks the SN location with an arrow.}
\end{center}
\end{figure*}

@@

As described below, we fit the SN \tomas\ spectrum in two steps.
First we determine a spectral classification -- and get a preliminary
estimate of the redshift and age -- using the SuperNova IDentification
(SNID) software \citep{Blondin:2007}.  Second, we refine the redshift
and age measurement using a custom Type Ia spectral template matching
program.

@@

The SNID program is designed to estimate the type, redshift, and age
of a SN spectrum through cross-correlation matching with a library of
template spectra.  To account for possible distortions in the broad
shape of the SN pseudo-continuum due to dust or instrumental
calibration effects, SNID divides each SED by a smooth cubic spline
fit. This effectively removes the shape of the SN pseudo-continuum to
leave behind a flat SED superimposed with spectral absorption and
emission features.  It is these features which drive the
cross-correlation fit, so the SNID approach is insensitive to the
overall color of the SED.  We used v2.0 of the SNID template library,
which includes template SEDs covering all Type Ia and Core Collapse
sub-classes, and has recently been updated with corrections and
improvements to the Type Ib/c templates \citep{Liu:2014}.

Using SNID to fit the \tomas\ spectrum, we find strong evidence that
the spectral type is a normal Type Ia, with $>$97\% of the matching
SEDs coming from Type Ia spectra.  
In SNID the goodness of fit is evaluated primarily through the {\it rlap} parameter, which measures the degree of wavelength overlap and the strength of the cross-correlation peak.  For SN \tomas\ the best matching Type Ia template is the normal Type Ia SN 2004fu at z=1.321 and age=2.4 days past peak, which delivers {\it rlap}=20.8.
Despite the relatively small rest-frame wavelength window of this spectrum, the SNID classification is quite definitive: \textcolor{red}{no \CCSN\ template achieves an {\it rlap} value better than 5.0 at any redshift.}

The best-fit Type Ia SEDs are also
consistent in redshift at $z=1.32\pm0.01$, and in phase at $0.4\pm1.7$
rest-frame days from peak (meaning MJD$_{\rm pk}$=56813$\pm$4).  At
this redshift, the absorption at $\sim$8700\AA\ corresponds to the
blended \ion{Ca}{2} H\&K features.  This \ion{Ca}{2} absorption is
commonly seen in Type Ia SN spectra near maximum light, although it is
also prominent in the spectra of Type Ib and Ic core collapse SNe
(\CCSNe).  The red color of the \tomas\ SED is qualitatively
consistent with a redshift of $z>1$ -- although this information was
not used by SNID for the template matching.  As we will see in
Section~\ref{sec:LightCurve}, this spectral classification of
SN \tomas\ is supported by the photometric information, which is also
most consistent with a Type Ia SN at $z=1.3$.

\subsection{Extended Type Ia Template Matching}
\label{sec:FittingWithAdditionalTypeIaTemplates}

@@

Relative to other SNe at $z>1$, the SN \tomas\ light curve was
unusually well sampled at rest-frame ultraviolet wavelengths, due to
the rapid cadence of the HFF imaging campaign. These ACS observations
therefore provide a tight constraint on the time of peak brightness
and the evolution of the SN color.  Supplemental observations with the
WFC3-IR camera provided critical rest-frame optical photometry,
enabling a measurement of the apparent luminosity distance through
light curve fitting.

\subsection{Photometric Classification}
\label{sec:PhotometricClassification}

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  This classification code uses the
{\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves at a range of redshifts and evaluate the classification
probability using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software \citep[SNANA,][]{Kessler:2009a}.  Due to
the expected impact of gravitational lensing magnification, we do not
include any prior on the intrinsic luminosity for any SN sub-class.

From this photometric analysis we derive a Type Ia classification
probability of $p(\rm Ia|{\bf \rm D})=XX\%$.  Marginalizing over the
light curve shape and color parameters of the SALT2 model, we extract
a posterior probability distribution in redshift that is well
described by a single Gaussian centered at $z=1.32\pm0.03$.
